Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-06 ind_files/03_field_eqns.tex
Make notation consistent

\frac{\partial p}{\partial t} &= \frac{1}{N_K(t)}\left[w(x|\psi)\psi(x,t) + \mu Q(x|\psi)\right] - \frac{\psi(x,t)}{N^2_K(t)}\int\limits_{\mathcal{T}}w(y|\psi)\psi(y,t) + \mu b^{\textrm{(mut)}}(y|\psi)dy\\
&= w(x|\psi)p(x,t) + \frac{\mu}{N_{K}(t)} Q(x|\psi) - p(x,t)\left(\int\limits_{\mathcal{T}}w(y|\psi)p(y,t)dy+\frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}} b^{\textrm{(mut)}}(y|\psi)dy\right)

@@

\setlength{\fboxsep}{2\fboxsep}\boxed{\frac{\partial p}{\partial t}(x,t) = \left[w(x|\psi) - \overline{w}(t)\right]p(x,t)+\frac{\mu}{N_K(t)}\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} b^{\textrm{(mut)}}(y|\psi)dy\right]}

@@

This is a continuous version of the replicator-mutator equation when each $x$ is viewed as a strategy. It also yields Kimura's continuum-of-alleles model when each $x$ is viewed as an allele, $Q(x|\psi)$ takes the form of a convolution of $\psi(x,t)$ with a mutation kernel, and the trait space is the entire real line, \emph{i.e.} $\mathcal{T} = \mathbb{R}$. To see this, let $b^{\textrm{(mut)}}(y|\psi) = \int\limits_{\mathbb{R}}m(y-z)\psi(z,t)dz$, where $m(x)$ is a mutation kernel, which by definition is normalized such that $\int\limits_{\mathbb{R}}m(x)dx = 1$. Let us further note that we have implicitly been assuming that the total number of individuals (scaled by $K$) remains finite at all times, \emph{i.e.} $N_K(t) = \int\limits_{\mathbb
R}\psi(x,t)dx < \infty \ \forall \ t$. Thus, $m(x)\psi(y,t) \in  \mathcal{L}^{1}(\mathbb{R}\times\mathbb{R}) \ \forall \ t$ and we can use the Fubini-Tonnelli theorem to interchange the order of integration of iterated integrals of $m(y-z)\psi(y)$. We are now ready to evaluate the rightmost integral of \eqref{cts_replicator_mutator}.\\

@@

\int\limits_{\mathbb{R}} b^{\textrm{(mut)}}(y|\psi)dy &= \int\limits_{\mathbb{R}}\int\limits_{\mathbb{R}} m(y-z)\psi(z,t)dzdy\nonumber\\

@@

where we have used the Fubini-Tonnelli theorem to go from the first step to the second, and have made the substitution $u = y-z$ to go from the third to the fourth step. We then note that since $m$ is a kernel, it satisfies $\int\limits_{\mathbb{R}} m(u)du = 1$, and \eqref{convolution_integral_intermediate} therefore becomes $\int\limits_{\mathbb{R}} b^{\textrm{(mut)}}(y|\psi)dy = N_K(t)$. Substituting this in \eqref{cts_replicator_mutator}, we have

@@

&+\frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}}f(x,t)\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} b^{\textrm{(mut)}}(y|\psi)dy\right]dx\nonumber\\
&= \overline{fw} - \overline{w}\cdot\overline{f} + \frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}}f(x,t)\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} b^{\textrm{(mut)}}(y|\psi)dy\right]dx\label{intermediate_for_cts_price}

@@

from which it is clear that we have obtained a version of the Price equation \eqref{nD_Price_time_dependent} for quantitative traits. For the special case $f(x)=x$, we have $\partial f/\partial t = 0$ and thus:

@@

In this case, some more familiar dynamics are recovered under the following additional assumptions:

>>>>

\frac{\partial p}{\partial t} &= \frac{1}{N_K(t)}\left[w(x|\psi)\psi(x,t) + \mu Q(x|\psi)\right] - \frac{\psi(x,t)}{N^2_K(t)}\int\limits_{\mathcal{T}}w(y|\psi)\psi(y,t) + \mu Q(y|\psi)dy\\
&= w(x|\psi)p(x,t) + \frac{\mu}{N_{K}(t)} Q(x|\psi) - p(x,t)\left(\int\limits_{\mathcal{T}}w(y|\psi)p(y,t)dy+\frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}} Q(y|\psi)dy\right)

@@

\setlength{\fboxsep}{2\fboxsep}\boxed{\frac{\partial p}{\partial t}(x,t) = \left[w(x|\psi) - \overline{w}(t)\right]p(x,t)+\frac{\mu}{N_K(t)}\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} Q(y|\psi)dy\right]}

@@

This is a continuous version of the replicator-mutator equation when each $x$ is viewed as a strategy. It also yields Kimura's continuum-of-alleles model when each $x$ is viewed as an allele, $Q(x|\psi)$ takes the form of a convolution of $\psi(x,t)$ with a mutation kernel, and the trait space is the entire real line, \emph{i.e.} $\mathcal{T} = \mathbb{R}$. To see this, let $Q(y|\psi) = \int\limits_{\mathbb{R}}m(y-z)\psi(z,t)dz$, where $m:\mathbb{R} \to [0,\infty)$ is a mutation kernel, which by definition is normalized such that $\int\limits_{\mathbb{R}}m(x)dx = 1$. Let us further note that we have implicitly been assuming that the total number of individuals (scaled by $K$) remains finite at all times, \emph{i.e.} $N_K(t) = \int\limits_{\mathbb
R}\psi(x,t)dx < \infty \ \forall \ t$. Thus, for any fixed $t > 0$, we have $m(y-z)\psi(z,t) \in  \mathcal{L}^{1}(\mathbb{R}\times\mathbb{R})$ and we can use the Fubini-Tonnelli theorem to interchange the order of integration of iterated integrals of $m(y-z)\psi(y)$. We are now ready to evaluate the rightmost integral of \eqref{cts_replicator_mutator}.\\

@@

\int\limits_{\mathbb{R}} Q(y|\psi)dy &= \int\limits_{\mathbb{R}}\int\limits_{\mathbb{R}} m(y-z)\psi(z,t)dzdy\nonumber\\

@@

where we have used the Fubini-Tonnelli theorem to go from the first step to the second, and have made the substitution $u = y-z$ to go from the third to the fourth step. We then note that since $m$ is a kernel, it satisfies $\int\limits_{\mathbb{R}} m(u)du = 1$, and \eqref{convolution_integral_intermediate} therefore becomes $\int\limits_{\mathbb{R}} Q(y|\psi)dy = N_K(t)$. Substituting this in \eqref{cts_replicator_mutator}, we have

@@

&+\frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}}f(x,t)\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} Q(y|\psi)dy\right]dx\nonumber\\
&= \overline{fw} - \overline{w}\cdot\overline{f} + \frac{\mu}{N_K(t)}\int\limits_{\mathcal{T}}f(x,t)\left[Q(x|\psi) - p(x,t)\int\limits_{\mathcal{T}} Q(y|\psi)dy\right]dx\label{intermediate_for_cts_price}

@@

from which it is clear that we have obtained a version of the Price equation \eqref{nD_Price_time_dependent} for quantitative traits (Note that \eqref{cts_price_general} is precisely the equation obtained by informally taking $m \to \infty$ in \eqref{nD_Price_time_dependent}). For the special case $f(x)=x$, we have $\partial f/\partial t = 0$ and thus:

@@

We can also recover some more familiar dynamics under the following additional assumptions:
